---
title: "Full airport scraper"
author: "Mafalda Gonzalez Gonzalez & Brad McKenzie"
date: "2025-03-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions:
This package will scrape information on the current status of flights in Spanish airports. It will use data from `aena.es` to scrape real time information on flights in the 2 hours prior and 12 hours ahead. AENA has information on 48 Spanish airports and heliports. You can return the status of each individual airport, or return the results for only the busiest airports. 

Ensure you have run the docker code as per the repository .readme instructions: 
```         
docker run -d -p 4445:4444 -p 5900:5900 --env VNC_NO_PASSWORD=1 --name selenium_firefox selenium/standalone-firefox-debug
```


# Which aiport?

In the github repos, you will find a table that has been generated from a separate script (`create airport code search table`). This joins AENA and wikipedia data to allow search for specific airport codes that match to the AENA website. 

While you can search with names of cities, some cities have multiple airports, so the 3 digit IATA code is the best search option. 

To input your airport code, update the airport_code below as a character string E.g. `airport_code <- "MAD"` would be used for Madrid Adolfo-Suarez Airport. We use "MAD" as the default search since we live here and are most interested.

```{r}
airport_code <- "EAS"

#create default search as MAD - Madrid, if no airport is selected above 
airport_code <- ifelse(airport_code == "", "MAD", airport_code)
```

#### Set your user-agent
(check your user agent [here](https://www.google.com/search?client=ubuntu&channel=fs&q=what%27s+my+user+agent&ie=utf-8&oe=utf-8))

We set this for ethical scraping, so that the page can recognise us as the ones downloading the data.

```{r, eval=FALSE}
# set_config(
#   user_agent("Mozilla/5.0 ....)
# )
```


### 1. Connect to Docker

Ensure Docker is downloaded and running on your device.

1️⃣ Docker command (image & ports) in PowerShell to start Selenium container

Selenium container with visible Firefox:

-   Optional =\> Install a VNC Viewer to see the browser:
    -   Connect to: localhost:5900
    -   No password required

```         
docker run -d -p 4445:4444 -p 5900:5900 --env VNC_NO_PASSWORD=1 --name selenium_firefox selenium/standalone-firefox-debug
```

Connection code:

```{r}
remDr <- remoteDriver(
  remoteServerAddr = "localhost",
  port = 4445L,
  browserName = "firefox"
)

remDr$open()
Sys.sleep(3) # pause for Docker to setup and run
```

We should now be connected to the remote Docker computer

## Load packages

Libraries
```{r}
library(RSelenium)
library(tidyverse)
library(rvest)
library(httr)

```

First, a quick check that the airport code is valid:
```{r}
airport_code_list <- read.csv("Airport_code_search_table.csv") |> 
  select(IATA_code) |> pull()

# if statement based on whether code exists
if (airport_code %in% airport_code_list) {
  # If good, YAY
  print(paste("Airport code", airport_code, "found. Let's get scraping!"))
} else {
  # Airport code is not found, return an error message
  stop("Airport code not found. \nPlease consult our 'Airport_code_search_table.csv' and choose a new IATA code")
}


```



## RSelenium scraping of AENA searchbar

#### Launch AENA flight info page. 

```{r}
remDr$navigate("https://www.aena.es/en/flight-info.html")
```

#### Click on the consent button

...well, we have selected reject cookies. 

```{r}
# select path and click

remDr$findElement(using = "xpath", "//div[@id = 'modal_footer']//button[@class='ecix-bold ecix-font-1-2em elix-accordion-btn-light elix-deny-all-btn']")$clickElement()

```

####Switch the arrivals and departures button

The default is to search in the arrivals page. We want to search airport by departures.

```{r}
# navigate and click change search button
remDr$findElement(using = "xpath", 
                  value = "//nav[@class='filtrovuelo in-page bg-primary']//div[@class='iconos']")$clickElement()

```

#### Click departures button

This just selects the box and allows us to input text to search:  

```{r}
# identify the departures box
departures_box_select <- remDr$findElement(using = "xpath", 
                                           value = "//nav[@class='filtrovuelo in-page bg-primary']//input[@id='Departuresin the Aena network:']")

# checking step ->  view the element in the current departures box placeholder and value.
departures_box_select$getElementAttribute("placeholder")
departures_box_select$getElementAttribute("value")

```

### Input the airport code to search 

To set airport of interest then choose from our dropdown menu. Airport code is used because it gives the most accurate response. 

```{r}

# send the airport code text to the departures box: 
#first, clear the box from any exisitng text (sometimes cookies retain names)
departures_box_select$clearElement()

#next enter the airport code
departures_box_select$sendKeysToElement(list(airport_code, key = "enter"))
```

The auto dropdown now appears with airports based on the search term. 

This step is just for those using a VNC viewer going step by step. You should see the dropdown values highlighted. 

```{r}
# testing step ->  highlight the autofill to see it's being used: 
remDr$findElement(using = "xpath",
                  value = "//div[@class='input aeropuertoAena autocompletable']//ul[@class='autoCompletable visible']")$highlightElement()
```

Select the dropdown option

This selects automatic based on the airport code. I have created this dynamically because sometimes multiple airports can appear when you search just the 3 letter airport code. Using the airport code only will ensure we click the exact airport we want. 

```{r}
# Find the node from the dropdown with our airport code, create dynamic xml link using sprintf()
airport_input <- sprintf("//span[. = '%s']", airport_code)

# Find and click the element
autofill_drop <- remDr$findElement(using = "xpath", value = airport_input)
autofill_drop$clickElement()

```

At this stage -> it should be a page that has filtered to your airport selected

#### Adjust time of search to next 12 hours

The timer defaults to around 2 hours before the current time on AENA. We read this information dynamically and set the +12 hour timer based on the start time.

First, we click the time box to enable the dropdown dynamic start time and end time information. 

```{r}

select_timebox <- remDr$findElement(using = "xpath", 
                  value = "//nav[@class='filtrovuelo in-page bg-primary']//section[@id='horario']")

select_timebox$clickElement()

```


### Impute start and end time of data collection

Now check we extract our start hour and start minute of the search. These are used to define our end values. 

```{r}
# extract the time from the dropdown
start_hr <- remDr$findElement(using = "xpath",
                  value = "//article[@class='custom-horas']//div[@class='desde hour_unit']//div[@class='valor hora']")

start_min <- remDr$findElement(using = "xpath",
                  value = "//article[@class='custom-horas']//div[@class='desde hour_unit']//div[@class='valor minuto']")

#extract the value as numeric
start_hr <- as.numeric(start_hr$getElementText())
start_min <- as.numeric(start_min$getElementText())

paste0("Our start time is ", start_hr, ":", start_min, " (24 clock used time). By default, this is usually between 2 and 4 hours from before the current time.")

```


Now we determine what our end time will be, 12 hours in advance. 

This will need to be a loop if the hour of the day is after 19 (as it will enter the next day) and then we will need to code the date selection variable too. As the inputs are characters. We calculate as numeric to determine end time.

```{r}
# create 12 hour ID to limit our search
end_hr_input <- case_when(start_hr<12 ~ start_hr+12, # for next 12 hrs if morning
                                       start_hr>12 ~ start_hr-12, # if afternoon, end will be morning
                                       TRUE ~ 0) # if start hr == 12, correct to 0 for midnight
end_min_input <- start_min # unchanged.

# show our end time parameter
paste0("Our end time will be ", end_hr_input, ":", end_min_input, " (24 clock used time)")

```

Input the end times to the drop down for the "to" column
The end unit uses the same xml but has "hasta" instead of "desde" in the a higher level node. We extract these nodes here:

```{r}
# select the end time options
# extract the time from the dropdown
end_hr <- remDr$findElement(using = "xpath",
                  value = "//article[@class='custom-horas']//div[@class='hasta hour_unit']//div[@class='valor hora']")

end_min <- remDr$findElement(using = "xpath",
                  value = "//article[@class='custom-horas']//div[@class='hasta hour_unit']//div[@class='valor minuto']")

```

And now we dynamically input our end times. 

Both the hour and minute slot require dynamic regex inputs as they both require 2 digit numeric values to input.

- If the hour time is below 10, we add an extra '0' to the search term. E.g. 7 becomes 07
- Similarly for the minutes slot, if the value is 0, we add a second zero (only options here are 0,15,30,45)

Input the end hour time 
- create dropdown within dynamic table 
- search for the correct time value dynamically in the dropdown
- click that exact node we find. 

```{r}
#select the hour button to launch hours dropdown
end_hr$clickElement()

# Construct dynamic XPath using sprintf() and regex for numeric input, ifelse corrects single digit codes and adds a zero in the regex. 
xpath_expression <- ifelse(end_hr_input<10, 
                           sprintf("//div[@class='valor hora']//ul[@tabindex='3']/li[text()='0%d']", as.numeric(end_hr_input)),
                           sprintf("//div[@class='valor hora']//ul[@tabindex='3']/li[text()='%d']", as.numeric(end_hr_input)))

# Find the element
element_hour <- remDr$findElement(using = "xpath", value = xpath_expression)

# Click the element (or perform other actions)
element_hour$clickElement()

```

Then dynamically select the minute slot:

This follows the same process as the end hour above. 

```{r}
#select the dropdown
end_min$clickElement()

# Find the path with our minute option in the dropdown
# again, if minute is zero, it adjusts to '0' but we need a 2 digit code to match the xml, use regex to add zero in
xpath_expression <- ifelse(end_min_input == 0, 
                        sprintf("//div[@class='valor minuto']//ul[@tabindex='4']/li[text()='0%d']", as.numeric(end_min_input)),
                        sprintf("//div[@class='valor minuto']//ul[@tabindex='4']/li[text()='%d']", as.numeric(end_min_input)))

# Find the element
element_min <- remDr$findElement(using = "xpath", value = xpath_expression)

# Click the minute option
element_min$clickElement()

```

Now click the ok button to hide the time dropdown.

```{r}
select_ok <- remDr$findElement(using = "xpath",
                  value= "//article[@class='custom-horas']//div[@class='botones']/button[@class='btn button-round-primary']")

#select ok to update end time
select_ok$clickElement()
```

#### Update calendar elements for date of search

Some dyanmic coding here is to enable a more defensible script again.

a) if the scraper is called in the morning, it will end on the same day.
b) it is after midday, the 12 hour search will end the following day. So the calendar selection must adjust accordingly.

First, we set our starting dates. This tracker is for current status, so it will always be today

```{r}
select_date_box <- remDr$findElement(using = "xpath",
                        value = "//section[@id='fechas']//div[@class='input']//input[@id='fecha']")
select_date_box$highlightElement()

select_date_box$clickElement()
```


Manual coding then will function:
a) base case: if day is non-last of month and it is before midday. Select the same day.
b) afternoon run: the end calendar day will be sometime the following morning and we need to factor this in.

```{r}
# set our date value for today
date_today <- Sys.Date() # in format yyyy-mm-dd
day_of_month <- as.numeric(day(date_today))

# set the xml paths to the numbered notes with regex, based on the day.
xpath_startcal_expression <- ifelse(as.numeric(day_of_month)<10, 
                                    # add leading zero to node name if <10
                                    sprintf("//div[contains(@class, 'react-datepicker__day react-datepicker__day--') and text()='%0d']", as.numeric(day_of_month)),
                                    # no leading zero if node 10+
                                    sprintf("//div[contains(@class, 'react-datepicker__day react-datepicker__day--') and text()='%d']", as.numeric(day_of_month)))

# set the expression for tomorrow to use if you run in the afternoon. This adds 1 day, so will only add zero if the day is 8 or earlier (on the 9th of the month the end date will be 10 from this call)
xpath_endcal_expression <- ifelse(as.numeric(day_of_month)+1<10, 
                                  sprintf("//div[contains(@class, 'react-datepicker__day react-datepicker__day--') and text()='%0d']", as.numeric(day_of_month+1)),
                                  sprintf("//div[contains(@class, 'react-datepicker__day react-datepicker__day--') and text()='%d']", as.numeric(day_of_month+1)))


# now we assign a value to each of these calendar input xml nodes 
cal_today_date <- remDr$findElement(using = "xpath", value = xpath_startcal_expression)
cal_tomorrow_date <- remDr$findElement(using = "xpath", value = xpath_endcal_expression)

```

Select either today or tomorrow based on the time of day. The default start date is the current date, so our selection automatically selects the end date. 

The calendar popout closes automatically after a click so we do not manually close. 

```{r}

if (start_hr < 12) {
  cal_today_date$clickElement()
} else {
  cal_tomorrow_date$clickElement()
}
```

Known error in the calendar search:

LAST DAY OF THE MONTH: If it is the last day of the month and after midday (so the 12 hour flight search ends on the first day of the next calendar month), there may be an error. I have tried to select the dates dynamically, but the only full date input is in long format e.g. "Monday, 21st of August, 2018" and R could not match this effectively with lubridate. So it was more robust to search by dynamically inputting the day of the month value and searching for that node. 

#### Press search to confirm our search options are input

This will also update the pages total number of flights upcoming. 

```{r}
search_button <- remDr$findElement(using = "xpath",
                        value = "//div[@class=' container'] //section[@id='boton']/button[@value='Search']")

search_button$clickElement()

```


#### Determine our active flight information
Read how many flights are upcoming:
Sometimes no flights exist so we need to return an error accordingly or skip.

```{r}

check_flights_exist <- function(remDr) {
  # First search: Number of flights
  tryCatch({
    flight_count_element <- remDr$findElement(
      using = "xpath",
      value = "//section[@id='infovuelos-info']//p[@class='h5 ligero']"
    )
    message("We have in-scope flights. Continuing.")
    return(invisible(NULL)) # End function if flights are found
  }, error = function(e) {
    # Check for the error message
    tryCatch({
      error_message_element <- remDr$findElement(
        using = "xpath",
        value = "//p[@class='denso' and text()='We cannot find any flights with the criteria given']"
      )
      stop("CANNOT COMPLETE: No flights found in scope for this airport with the given criteria.") # Stop with error message
    }, error = function(e) {
      # If neither is found, re-throw the original error
      stop("Unexpected error occurred: ", e$message)
    })
  })
}

check_flights_exist(remDr=remDr)
```

```{r, eval=F}
number_flights_upcoming_selected <- 
  remDr$findElement(using = "xpath",
                    value = "//section[@id='infovuelos-info']//p[@class='h5 ligero']")

# paste our search terms
search_details <- number_flights_upcoming_selected$getElementText()
search_details

# extract the number of flights
number_flights_searched <- as.numeric(substr(search_details, 1,4))
number_flights_searched

```

The page returns 20 rows originally and when you click "see more" it adds 20 more. We specify the number of clicks based on our in scope searches and add a sleep in between. 
This needs to be an exact number, because the "see more" button disappears once you read the end and we don't want to produce an error:

If there are 95 flights, we would click 4 times. We start with 20, then search 4 more times to return up to 100 results. 

If there are 1555 flights, we would need to click 77 times. We add a sleep in between to avoid being banned and overloading the website. 

```{r}
num_clicks <- ceiling((number_flights_searched-20)/ 20)
paste("We will require",num_clicks,"clicks. This will take about 1 second per click")

```

We create a function to through the "see more" function to expand and view all flights. Adding a sleep to not overload the website does extend the wait for larger airports. It is generally under 2 minutes run time though. 

```{r}
multi_click_viewmore <- function(remDr) {
  while (TRUE) { # Loop indefinitely until can't find the "see more" button
    tryCatch({
      click_button <- remDr$findElement(using = "xpath", value = "//section[@id='infovuelos-tabla']//p[@class='btnIconText btn-see-more']//span[@class='icon icon-Mas_T']")
      click_button$clickElement()
      Sys.sleep(1)
      # based on the error type, we stop the loop and return message.
    }, error = function(e) {
      if (grepl("Unable to locate element", e$message)) {
        # Element not found, stop the loop
        message("Button not found. Stopping loop.")
        return(invisible(NULL)) # Exit the function
      } else {
        # Other error
        stop(e)
      }
    })
  }
}
```

Now we run this click more to expand our page to view all flights

```{r, eval=FALSE}
multi_click_viewmore(remDr = remDr)
```

#### Now we extract our page html file for analysis.

```{r}
# 1. Get the page source (updated HTML)
aena_page_source <- remDr$getPageSource()[[1]]

# 2. Parse the HTML using rvest
aena_flight_information <- read_html(aena_page_source)
```

We now have our html script parsed into the R local program. 

# Nodes scraping

## Loading functions

### Flight information function 

```{r function flight info}
categorise_flight <- function(flight) {
  
  clean_text <- flight %>% 
    html_text(trim = TRUE)
  
  departure_time_current <- clean_text %>% 
    .[[1]] %>% 
    str_sub(., 1, 5)
  
  departure_time_original <- clean_text %>% 
    .[[1]] %>% 
    str_sub(., -5, -1)


  flight_data <- tibble(
    departure_time_current = departure_time_current,
    departure_time_original = departure_time_original,
    flight_number = clean_text[[2]],
    company = clean_text[[3]],
    destination = clean_text[[4]],
    terminal = clean_text[[5]],
    check_in_desk = clean_text[[6]],
    boarding_gate = clean_text[[7]],
    getting_there = clean_text[[8]],
    flight_status = clean_text[[9]]
  )
  
  return(flight_data)
}

```

### Iteration through different flights (same Airport) function 

```{r function scrape_flights}
scrape_flights <- function(url) {

  all_flight_nodes <- url %>% 
    html_nodes(xpath = "//div[@class='fila']//div[@class='linea micro']")
  
  flights_data <- map_dfr(all_flight_nodes, function(flight) {
    
    children <- html_children(flight)  
    categorise_flight(children)        
  })
  
  return(flights_data)
}

```

### Delay time and status function 

```{r function categorise_delay}

categorise_delay <- function(flights_data) {
  
  flights_data %>%
    mutate(
      
      # conversion to datetime
      departure_time_current = today() + hm(departure_time_current),  
      departure_time_original = today() + hm(departure_time_original),
      
      # midnight crossover fix: adjust if current time is >30 min earlier than original
      departure_time_current = case_when(
        departure_time_current < departure_time_original & 
          (departure_time_original - departure_time_current) > minutes(30) 
        ~ departure_time_current + days(1),
        TRUE ~ departure_time_current
      ),
      
      # delay in minutes
      delay_minutes = as.numeric(difftime(departure_time_current, departure_time_original, units = "mins")),
      
      #  delay status
      delay_status = case_when(
        delay_minutes >= -5 & delay_minutes <= 10 ~ "On Time",
        delay_minutes > 10 ~ "Late", 
        delay_minutes < -5 ~ "Early"
      )
    )
}

```

### Unique flights function

```{r function group flights}
group_flights <- function(flights_data) {
  flights_data %>%
    group_by(departure_time_current, destination, boarding_gate) %>%
    summarise(
      departure_time_original = first(departure_time_original),
      flight_status = first(flight_status),
      delay_minutes = first(delay_minutes),
      delay_status = first(delay_status),
      destination = first(destination),
      flight_number = str_c(unique(flight_number), collapse = ", "),
      company = str_c(unique(company), collapse = ", "),
      check_in_desk = first(check_in_desk),
      terminal = first(terminal),
      boarding_gate = first(boarding_gate),
      getting_there = first(getting_there),
      .groups = "drop"
    ) %>%
  select(departure_time_current, departure_time_original,
         flight_status, delay_minutes, delay_status,
         destination, flight_number, company, check_in_desk,
         terminal, boarding_gate, getting_there)
}


```

## MOTHERFUNCTION 

```{r motherfunction}

scrape_flights <- function(url) {
  
  # all flight nodes
  all_flight_nodes <- url %>% 
    html_nodes(xpath = "//div[@class='fila']//div[@class='linea micro']")
  
  # no flights = empty tibble
  if (length(all_flight_nodes) == 0) {
    return(tibble())
  }

  # child elements from each flight
  flights_data <- map_dfr(all_flight_nodes, function(flight) {
    
    flight <- html_children(flight)
    
  # categorise_flight function
    categorise_flight(flight)        
  })
  
  # categorise_delay function 
  flights_data <- categorise_delay(flights_data)
  
  # gropu_flights function 
  flights_data <- group_flights(flights_data)

  return(flights_data)
}

```


```{r}
scrape_flights(aena_flight_information)

```

# CSV 

TODO: 
```{r, eval=F}
# -------------------------------------------------------
airport_urls <- list(
  "MAD" = "Flight information _ Infovuelos _ Aena.html", 
  "BCN" = "Información sobre vuelos _ Infovuelos _ Aena_BCN.html")
```


## individual airports

loop through multiple airport URLs + scrape the flight data + save each tibble as a CSV file with the airport name in the filename

```{r function CSV individual airports}
tibble_flights <- list()

save_flights_data <- function(url, airport_name) {
  
  # scrape flight data
  tibble_flights <- scrape_flights(url)
  
  # no data = skip saving + message
  if (nrow(tibble_flights) == 0) {
    
    print(paste("No flights found for", 
                airport_name, 
                "- Skipping CSV save."))
    return(NULL)
    
  }

  # timestamp column
  tibble_flights$date_saved <- format(Sys.time(), "%Y-%m-%d %H:%M")

  # file path (save in the current working directory) with airport name
  folder_path <- "flights_data"
  file_path <- paste0(folder_path, "/", airport_name, "_flights.csv")

  # make sure the directory exists
  if (!dir.exists(folder_path)) {
    dir.create(folder_path)
  }
  
  # silent try
  response <- try(read_csv(file_path, show_col_types = FALSE), silent = TRUE)

  # if the file doesnt exist => create it 
  
  if (inherits(response, "try-error")) {
    
    print(paste("Creating new file for", airport_name))
    write_csv(tibble_flights, file_path)
    
  } else {
    
    # if file exists => append new data
    
    print(paste0("Updating ", airport_name, "_flights.csv file"))
    
    rbind(response, tibble_flights) %>% write_csv(file_path)
    
  }
}
```

```{r}
save_flights_data(aena_flight_information, "MAD")
```


## LOOP for ALL INDIVIDUAL airports

```{r command CSV loop all individual airports}
# loop through all airports and save data
walk2(airport_urls, names(airport_urls), save_flights_data)

```

## all airports (we combine all datasets): 

```{r function combine_flights_data}
combine_flights_data <- function(folder_path = "flights_data") {
  
  # get all csv files
  csv_files <- list.files(folder_path, pattern = "\\.csv$", full.names = TRUE)
  
  # no files no info
  if (length(csv_files) == 0) {
    print("No CSV files found in the folder.")
    return(NULL)
  }
  
  # all CSVs + map_dfr to avoid looping (loops suck)! 
  all_data <- map_dfr(csv_files, function(file) {
    
    data <- read_csv(file, show_col_types = FALSE)
    
    # extract airport name = filename without extension
    airport_name <- tools::file_path_sans_ext(basename(file))
    
    # add airport column 
    data$source <- airport_name
    
    return(data)
  })
  
  return(all_data)
}
```

RESULT: FUNCTIONED

```{r function CSV save_combined_flights}
save_combined_flights <- function(folder_path = "flights_data", output_file = "flights_data/all_flights_combined.csv") {
  
  # Combine all airport data
  all_airports_data <- combine_flights_data(folder_path)
  
  # no data = no info
  if (nrow(all_airports_data) == 0) {
    
    print("No data to save.")
    return(NULL)
    
  }

  # silent try
  res <- try(read_csv(output_file, show_col_types = FALSE), silent = TRUE)

  # same as before
  if (inherits(res, "try-error")) {
    
    print("Creating new all_flights_combined.csv file")
    write_csv(all_airports_data, output_file)
    
  } else {
    
    print("Updating all_flights_combined.csv file")
    
    rbind(res, all_airports_data) %>% write_csv(output_file)
    
  }
}
```


```{r}
save_combined_flights()

```

