---
title: "Plotting and summary stats"
author: "Mafalda Gonzalez Gonzalez & Bradley McKenzie"
date: "2025-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a simple script to summarise and plot the output from the AENA scraper. 

```{r}
library(tidyverse)
library(ggplot2)
```

### TO DO: 
Update the location of your saved .csv from the scraper `Airport-scraper-full.Rmd"

Then you can run this full script, it will create a basic summary output and aver

```{r}
# link your output below. May take form like: 
#airport_data <- read_csv("flights_data/MAD_flights.csv")

airport_data <- read_csv("")
```
## View your table
```{r}
airport_data

#if you want to view delays only
```



## Summary statistics

```{r}
# summary table output
summary_stats <- airport_data |> 
  summarise(
    `Number of flights` = n(),
    `Flights departed` = sum(flight_status == "Ended", na.rm=T),
    `Flights soon to depart` = sum(flight_status %in% c("Closed", "Boarding", "Delayed"), na.rm=T),
    `AENA reported delays` = sum(flight_status =="Delayed", na.rm=T), # their reported delays
    `Scraper delayed flights` = sum(delay_status == "Late", na.rm=T),
    `Average minutes per scraper delay` = round(mean(delay_minutes[delay_status == "Late"], na.rm = TRUE),2),
    `Scraper on time flights` = sum(delay_status == "On Time", na.rm=T),
    `Scraper early flights` = sum(delay_status == "Early", na.rm=T)) |> 
  pivot_longer(cols = everything(),
               names_to = "variable",
               values_to = "value")

summary_stats

```

## Plot current airport delays 

This will work for a table with all flight information. Uncomment the line `facet_wrap(~airport_names)+` if you have the joined dataframe for each airport and want create a plot for each airport

Plot the distribution of upcoming flights based on the data returned from your scraper. 

```{r}

plot_upcoming_flights <- airport_data |>
  # ensure our setting date variable is formatted for the X-axis 
  mutate(departure_time_original = as.POSIXct(departure_time_original, 
                                              format = "%Y-%m-%d %H:%M:%S")) |>
  
  ggplot(aes(x = departure_time_original, 
             y = delay_minutes, 
             size = abs(delay_minutes))) + 
  geom_point() +
  geom_smooth(method="lm")+
  annotate("rect", xmin = min(test_flights$departure_time_original), 
           xmax = max(test_flights$departure_time_original), 
           ymin = -10, ymax = 10, fill = "green", alpha = 0.3) +
  annotate("rect", xmin = min(test_flights$departure_time_original), 
           xmax = max(test_flights$departure_time_original), 
           ymin = 10, ymax = Inf, fill = "red", alpha = 0.3) +
  annotate("rect", xmin = min(test_flights$departure_time_original), 
           xmax = max(test_flights$departure_time_original), 
           ymin = -Inf, ymax = -10, fill = "orange", alpha = 0.3) +
  theme_minimal()+
#  facet_wrap(~airport_names)+
  labs(title = "Current flight information by delay time", 
       x = "Delay per flight in minutes", 
       y = "Original departure time",
       size = "Delay length (minutes)")

```

