---
title: "scrapie project"
author: "Mafalda González González"
format: 
  html:
    embed-resources: true
editor: visual
---

# Library

```{r}

library(xml2)
library(httr)
library(httr2)
library(tidyverse)
library(rvest)
library(dplyr)
library(stringr)
library(purrr)
library(lubridate)
library(RSelenium)
```

# Airport

```{r}
madrid <- "https://www.aena.es/en/flight-info.html"
url <- madrid

library(httr)
url <- GET("https://www.aena.es/en/flight-info.html", add_headers(`User-Agent` = "Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/131.0.0.0 Safari/537.36; Mafalda Gonzalez Gonzalez / mafalda.g.gonzalez@gmail.com"))

set_config(
  user_agent("Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/131.0.0.0 Safari/537.36; Mafalda Gonzalez Gonzalez / mafalda.g.gonzalez@gmail.com")
)
```

# Scrápie

problem: page is loaded dynamically with javascript AFTER loading the sourv

-   page source = initial html send by the server
-   inspect element = fully loaded website =\> including dynamically added content with java

ALSO: flight delays and info is loaded with API requests! not static content!

this wont work:

```{r}
#theoretically code if website was static
url %>% 
  read_html() %>% 
  xml_find_all("//div[@class='hora']") %>% 
  xml_text()

# can only go one node lower than the root
url %>% 
  read_html() %>% 
  html_nodes(xpath = "//body/div[@id='infovuelos']") %>% 
  html_structure()

# finds only horario, as is in the html source code, but not hora. horario is actually not anymore in the final output script 
url %>% 
  read_html() %>% 
  html_nodes(xpath = "//div[contains(@class, 'hora')]") 
```

## with downloaded page

```{r}
url <- "Flight information _ Infovuelos _ Aena.html"
```

# Select flights

we are looking to select each individual fila info so that we can download and safe the data

different options

## with CSS

```{r}

url %>% 
  read_html() %>% 
  html_nodes(xpath = "//div[@class='fila']") %>% 
  .[[1]] %>% 
  html_nodes(".linea.micro") %>% # CSS selector instead of xpath => xpath 
  html_children()
```

=\> works

-   we find the filas (xml_find_all or html_nodes would be the same here)
-   we find the first element
-   we use **CSS** instead of xpath with hmtl_nodes() so that we select within out first element selected! =\> we then find the root of all of the information of that fila - hmtl_children to get all the individual information into different outputs

## with XPath

next try:

```{r}
one_flight <- url %>% 
  read_html() %>% 
  html_nodes(xpath = "//div[@class='fila']//div[@class='linea micro']") %>% 
  .[[1]] %>% 
  html_children()

one_flight
```

same as before but we select the linea micro first, and then out of all linea micros we pick the first one. same output different strategy. idk whats better or worse. ask jorge?

# Information into data bases

```{r}

clean_text <- one_flight %>% 
  html_text(trim = TRUE)

clean_text

data <- tibble(
  departure_time = clean_text[[1]],
  flight_number = clean_text[[2]],
  company = clean_text[[3]],
  destination = clean_text[[4]],
  terminal = clean_text[[5]],
  check_in_desk = clean_text[[6]],
  boarding_gate = clean_text[[7]],
  getting_there = clean_text[[8]],
  flight_status = clean_text[[9]]
)

data
```

problem: departure time:

if departure time has more than 5 characters: split it into original departure time and late_departure_time

-   either with CSS (XPATH DOES NOT WORK: if we used xml_find_all or html_nodes(xpath = "//span\[not(contains( \@class, 'tachado'))\]") =\> we would go back into the source code, and we just want to saty inside one_flight\[\[1\]\])

-   or with text manipulation/regex

```{r}

one_flight

# data$departure_time_current <- 

# CSS
one_flight[[1]] %>% 
  html_nodes("span:not(.tachado)") %>% 
  html_text(trim = TRUE)

# regex 
clean_text <- one_flight %>% 
  html_text(trim = TRUE)

clean_text %>% 
  .[[1]] %>% 
  str_sub(., 1, 5)



# data$departure_time_original <- 

# CSS
one_flight[[1]] %>%
  html_nodes("span.tachado") %>%
  html_text(trim = TRUE) %>% 
  str_c(collapse = " ")

# function: ifelse(length(.) > 0, ., departure_time_current)

# regex
clean_text %>% 
  .[[1]] %>% 
  str_sub(., -5, -1)


```

=\> better with str_sub(., 1, 5) caus eit will automatically put the same time if there is no "tachado"

## f: flight information

```{r}
categorise_flight <- function(flight) {
  
  clean_text <- flight %>% 
    html_text(trim = TRUE)
  
  departure_time_current <- clean_text %>% 
    .[[1]] %>% 
    str_sub(., 1, 5)
  
  departure_time_original <- clean_text %>% 
    .[[1]] %>% 
    str_sub(., -5, -1)


  flight_data <- tibble(
    departure_time_current = departure_time_current,
    departure_time_original = departure_time_original,
    flight_number = clean_text[[2]],
    company = clean_text[[3]],
    destination = clean_text[[4]],
    terminal = clean_text[[5]],
    check_in_desk = clean_text[[6]],
    boarding_gate = clean_text[[7]],
    getting_there = clean_text[[8]],
    flight_status = clean_text[[9]]
  )
  
  return(flight_data)
}

categorise_flight(one_flight)
```

# Read all flights

```{r}
library(purrr)

# all flight nodes (each div.linea.micro is a flight)
all_flight_nodes <- url %>% 
  read_html() %>% 
  html_nodes(xpath = "//div[@class='fila']//div[@class='linea micro']")


```

map() to loop through all flights in the xml_nodeset

```{r}
# applying html_children() to each flight node and then categorise_flight()
flights_data <- map_dfr(all_flight_nodes, function(flight) {
  flight <- html_children(flight)  # extract individual flight details
  categorise_flight(flight)        
})

flights_data

```

## f: ITERATE through flights

```{r}
scrape_flights <- function(url) {

  all_flight_nodes <- url %>% 
    read_html() %>% 
    html_nodes(xpath = "//div[@class='fila']//div[@class='linea micro']")
  
  flights_data <- map_dfr(all_flight_nodes, function(flight) {
    children <- html_children(flight)  # Extract individual flight details
    categorise_flight(children)        # Process the extracted children
  })
  
  return(flights_data)
}

scrape_flights(url)

flights_data <- scrape_flights(url)

```

# Delay classes with lubridate

-   LUBRIDATE: convert times into proper hms (hour-minute-second) format
-   cases where flights cross over into the next day e.g., 23:40 to 00:10 should be recognized as a delay
-   =\> calculate time difference correctly

```{r}
# test 
x <- tibble(
  departure_time_current = c("00:10", "14:30", "12:00", "23:59", "06:00"),
  departure_time_original = c("23:40", "14:30", "12:30", "23:00", "06:31")
)

library(lubridate)

x %>%
  mutate(
    # lubridate conversion into datetime: today + hour-minute
    departure_time_current = today() + hm(departure_time_current),  
    departure_time_original = today() + hm(departure_time_original),
    
    # midnight flights crossover fix  
    departure_time_current = case_when(
      
      # if current time is smaller than original & diff between them is more than 30minutes (we assume the max time a flight can be early is 30minutes) => we assume current is the next day
      
      # => if its smaller => if its 00:10 and the original was 23:10 then we want to do 24:10>23:10, for that we add the 24hours
      
      departure_time_current < departure_time_original & (departure_time_original - departure_time_current) > minutes(30) 
      ~ departure_time_current + days(1),
      TRUE ~ departure_time_current
    ),
    
    # delay in minutes
    delay_minutes = as.numeric(difftime(departure_time_current, departure_time_original, units = "mins")),
    
    # delay status
    delay_status = case_when(
      delay_minutes >= -5 & delay_minutes <= 10 ~ "On Time", # taking absolute value so that if its smaller than ten minutes early or late, its still on time theoretically 
      delay_minutes > 10 ~ "Late", 
      delay_minutes < -5 ~ "Early"
    )
  )


```

## f: delay time and status

```{r}
categorise_delay <- function(flight) {
  
  flight %>%
    mutate(
      
      # conversion to datetime
      departure_time_current = today() + hm(departure_time_current),  
      departure_time_original = today() + hm(departure_time_original),
      
      # midnight crossover fix: adjust if current time is >30 min earlier than original
      departure_time_current = case_when(
        departure_time_current < departure_time_original & 
          (departure_time_original - departure_time_current) > minutes(30) 
        ~ departure_time_current + days(1),
        TRUE ~ departure_time_current
      ),
      
      # delay in minutes
      delay_minutes = as.numeric(difftime(departure_time_current, departure_time_original, units = "mins")),
      
      #  delay status
      delay_status = case_when(
        delay_minutes >= -5 & delay_minutes <= 10 ~ "On Time",
        delay_minutes > 10 ~ "Late", 
        delay_minutes < -5 ~ "Early"
      )
    )
}

categorise_delay(flights_data)

flights_data <- categorise_delay(flights_data)
```

# unique flights

```{r, eval=F}
flights_data
```

```{r}
flights_data %>%
    group_by(departure_time_current, destination, boarding_gate) %>%
    summarise(
      departure_time_original = first(departure_time_original),
      flight_status = first(flight_status),
      delay_minutes = first(delay_minutes),
      delay_status = first(delay_status),
      destination = first(destination),
      flight_number = str_c(unique(flight_number), collapse = ", "),
      company = str_c(unique(company), collapse = ", "),
      check_in_desk = first(check_in_desk),
      terminal = first(terminal),
      boarding_gate = first(boarding_gate),
      getting_there = first(getting_there),
      .groups = "drop"
    ) %>%
  # order of the columns displayed
  select(departure_time_current, departure_time_original,
         flight_status, delay_minutes, delay_status,
         destination, flight_number, company, check_in_desk,
         terminal, boarding_gate, getting_there)
```

## f: unique flights

```{r}
group_flights <- function(flights_data) {
  flights_data %>%
    group_by(departure_time_current, destination, boarding_gate) %>%
    summarise(
      departure_time_original = first(departure_time_original),
      flight_status = first(flight_status),
      delay_minutes = first(delay_minutes),
      delay_status = first(delay_status),
      destination = first(destination),
      flight_number = str_c(unique(flight_number), collapse = ", "),
      company = str_c(unique(company), collapse = ", "),
      check_in_desk = first(check_in_desk),
      terminal = first(terminal),
      boarding_gate = first(boarding_gate),
      getting_there = first(getting_there),
      .groups = "drop"
    ) %>%
  select(departure_time_current, departure_time_original,
         flight_status, delay_minutes, delay_status,
         destination, flight_number, company, check_in_desk,
         terminal, boarding_gate, getting_there)
}

group_flights(flights_data)
  
```

# FUNCTION

```{r}

scrape_flights <- function(url) {
  
  # all flight nodes
  all_flight_nodes <- url %>% 
    read_html() %>% 
    html_nodes(xpath = "//div[@class='fila']//div[@class='linea micro']")
  
  # no flights = empty tibble
  if (length(all_flight_nodes) == 0) {
    return(tibble())
  }

  # child elements from each flight
  flights_data <- map_dfr(all_flight_nodes, function(flight) {
    
    flight <- html_children(flight)
    
  # categorise_flight function
    categorise_flight(flight)        
  })
  
  # categorise_delay function 
  flights_data <- categorise_delay(flights_data)
  
  # gropu_flights function 
  flights_data <- group_flights(flights_data)

  return(flights_data)
}

tibble_flights <- scrape_flights(url)

```

# f: CSV

## individual airports

loop through multiple airport URLs + scrape the flight data + save each tibble as a CSV file with the airport name in the filename

```{r}

tibble_flights <- list()

save_flights_data <- function(url, airport_name) {
  
  # scrape flight data
  tibble_flights <- scrape_flights(url)
  
  # no data = skip saving + message
  if (nrow(tibble_flights) == 0) {
    
    print(paste("No flights found for", 
                airport_name, 
                "- Skipping CSV save."))
    return(NULL)
    
  }

  # timestamp column
  tibble_flights$date_saved <- format(Sys.time(), "%Y-%m-%d %H:%M")

  # file path (save in the current working directory) with airport name
  folder_path <- "flights_data"
  file_path <- paste0(folder_path, "/", airport_name, "_flights.csv")

  # make sure the directory exists
  if (!dir.exists(folder_path)) {
    dir.create(folder_path)
  }
  
  # silent try
  response <- try(read_csv(file_path, show_col_types = FALSE), silent = TRUE)

  # if the file doesnt exist => create it 
  
  if (inherits(response, "try-error")) {
    
    print(paste("Creating new file for", airport_name))
    write_csv(tibble_flights, file_path)
    
  } else {
    
    # if file exists => append new data
    
    print(paste0("Updating ", airport_name, "_flights.csv file"))
    
    rbind(response, tibble_flights) %>% write_csv(file_path)
    
  }
}

# -------------------------------------------------------
airport_urls <- list(
  "MAD" = "Flight information _ Infovuelos _ Aena.html", 
  "BCN" = "Información sobre vuelos _ Infovuelos _ Aena_BCN.html")

# loop through all airports and save data
walk2(airport_urls, names(airport_urls), save_flights_data)

```

## all airports (we combine all datasets):

```{r}
combine_flights_data <- function(folder_path = "flights_data") {
  
  # get all csv files
  csv_files <- list.files(folder_path, pattern = "\\.csv$", full.names = TRUE)
  
  # no files no info
  if (length(csv_files) == 0) {
    print("No CSV files found in the folder.")
    return(NULL)
  }
  
  # all CSVs + map_dfr to avoid looping (loops suck)! 
  all_data <- map_dfr(csv_files, function(file) {
    
    data <- read_csv(file, show_col_types = FALSE)
    
    # extract airport name = filename without extension
    airport_name <- tools::file_path_sans_ext(basename(file))
    
    # add airport column 
    data$source <- airport_name
    
    return(data)
  })
  
  return(all_data)
}
```

RESULT: MANUAL

```{r, eval=F}
all_airports_data <- combine_flights_data()

print(all_airports_data)

write_csv(all_airports_data, "flights_data/all_flights_combined.csv")
```

RESULT: FUNCTIONED

```{r}
save_combined_flights <- function(folder_path = "flights_data", output_file = "flights_data/all_flights_combined.csv") {
  
  # Combine all airport data
  all_airports_data <- combine_flights_data(folder_path)
  
  # no data = no info
  if (nrow(all_airports_data) == 0) {
    
    print("No data to save.")
    return(NULL)
    
  }

  # silent try
  res <- try(read_csv(output_file, show_col_types = FALSE), silent = TRUE)

  # same as before
  if (inherits(res, "try-error")) {
    
    print("Creating new all_flights_combined.csv file")
    write_csv(all_airports_data, output_file)
    
  } else {
    
    print("Updating all_flights_combined.csv file")
    
    rbind(res, all_airports_data) %>% write_csv(output_file)
    
  }
}

save_combined_flights()

```

# Selenium test

```{r}
library(RSelenium)


remDr <- remoteDriver(
  remoteServerAddr = "localhost",
  port = 4445L,
  browserName = "firefox"
)

remDr$open()

remDr$navigate("https://www.google.com")


```

🎉 Amazing! 🎉 I did it! 🚀

For Prof:

1️⃣ Docker command (image & ports) in PowerShell to start Selenium container

Selenium container with visible Firefox:

-   Optional =\> Install a VNC Viewer to see the browser:
    -   Connect to: localhost:5900
    -   No password required

```         
docker run -d -p 4445:4444 -p 5900:5900 --env VNC_NO_PASSWORD=1 --name selenium_firefox selenium/standalone-firefox-debug
```

RSelenium connection code:

```{r}
remDr <- remoteDriver(
  remoteServerAddr = "localhost",
  port = 4445L,
  browserName = "firefox"
)

remDr$open()
```

# Selenium functions

# FUNctionastic

Hence, having done tested of the above for an individual airport, we can create functions to do the following things for a list of airports:

-   Navigate to correct page and search for flights
-   Select settings / parameters: airport, time, date, search flights
-   Expand flights list
-   Extract and save HTML for each Airport on the list

We will do the functions below. While they have no explanations, as we have gone over the idea of the code above, they do include comments to remember what each command does.

## Navigate to correct page and search for flights

```{r}
search_flights_aena <- function(airport_code) {
  
  # navigate to AENA flight page
  remDr$navigate("https://www.aena.es/en/flight-info.html")
  Sys.sleep(2)

  # reject cookies
  remDr$findElement(
    using = "xpath", 
    "//div[@id = 'modal_footer']//button[@class='ecix-bold ecix-font-1-2em elix-accordion-btn-light elix-deny-all-btn']")$clickElement()
  Sys.sleep(2)

  # switch to departures tab
  remDr$findElement(using = "xpath", 
                    "//nav[@class='filtrovuelo in-page bg-primary']//div[@class='iconos']")$clickElement()
  Sys.sleep(2)

  # select departures input box
  departures_box_select <- remDr$findElement(
    using = "xpath",
    "//nav[@class='filtrovuelo in-page bg-primary']//input[@id='Departuresin the Aena network:']")
  departures_box_select$clearElement()
  departures_box_select$sendKeysToElement(list(airport_code, key = "enter"))
  Sys.sleep(2)

  # select dropdown result
  airport_input <- sprintf("//span[. = '%s']", airport_code)
  remDr$findElement(using = "xpath", value = airport_input)$clickElement()
  Sys.sleep(2)
  
}

search_flights_aena(MAD)

```

## Select settings for time, date, Airport and search for flights

```{r}
set_time_and_date <- function() {
  
  # open time settings
  remDr$findElement(
    using = "xpath",
    "//nav[@class='filtrovuelo in-page bg-primary']//section[@id='horario']")$clickElement()
  Sys.sleep(2)

  # get current time
  start_hr <- as.numeric(
    remDr$findElement(
      using = "xpath",
      "//article[@class='custom-horas']//div[@class='desde hour_unit']//div[@class='valor hora']")$getElementText())
  start_min <- as.numeric(
    remDr$findElement(
      using = "xpath",
      "//article[@class='custom-horas']//div[@class='desde hour_unit']//div[@class='valor minuto']")$getElementText())

  end_hr_input <- ifelse(start_hr < 12, start_hr + 12, start_hr - 12)
  end_min_input <- start_min

  # select end hour
  remDr$findElement(
    using = "xpath",
    "//article[@class='custom-horas']//div[@class='hasta hour_unit']//div[@class='valor hora']")$clickElement()
  xpath_hr <- sprintf("//div[@class='valor hora']//ul[@tabindex='3']/li[text()='%02d']", end_hr_input)
  remDr$findElement(
    using = "xpath", value = xpath_hr)$clickElement()
  Sys.sleep(1)

  # select end minute
  remDr$findElement(
    using = "xpath",
    "//article[@class='custom-horas']//div[@class='hasta hour_unit']//div[@class='valor minuto']")$clickElement()
  xpath_min <- sprintf("//div[@class='valor minuto']//ul[@tabindex='4']/li[text()='%02d']", end_min_input)
  remDr$findElement(
    using = "xpath", value = xpath_min)$clickElement()
  Sys.sleep(1)

  # confirm time selection
  remDr$findElement(
    using = "xpath",
    "//article[@class='custom-horas']//div[@class='botones']/button[@class='btn button-round-primary']")$clickElement()
  Sys.sleep(2)

  # Adjust date if necessary
  date_today <- Sys.Date()
  day_of_month <- as.numeric(format(date_today, "%d"))

  xpath_date <- sprintf("//div[contains(@class, 'react-datepicker__day') and text()='%d']", day_of_month)
  if (start_hr >= 12) {
    xpath_date <- sprintf("//div[contains(@class, 'react-datepicker__day') and text()='%d']", day_of_month + 1)
  }

  remDr$findElement(
    using = "xpath", value = xpath_date)$clickElement()
  Sys.sleep(2)
}

```

## Expand flights list

Multiclick function of before:

```{r}
multi_click_viewmore <- function(remDr) {
  
  while (TRUE) { # Loop indefinitely until can't find the "see more" button
    
    tryCatch({
      
      click_button <- remDr$findElement(using = "xpath", value = "//section[@id='infovuelos-tabla']//p[@class='btnIconText btn-see-more']//span[@class='icon icon-Mas_T']")
      click_button$clickElement()
      Sys.sleep(1)
      # based on the error type, we stop the loop and return message.
      
    }, error = function(e) {
      
      if (grepl("Unable to locate element", e$message)) {
        
        # element not found, stop the loop
        message("Button not found. Stopping loop.")
        return(invisible(NULL)) # exit the function
        
      } else {
        
        # other error
        stop(e)
      }
    })
  }
}
```

```{r}
expand_flights_list <- function() {
  
  # get total number of flights
  search_details <- remDr$findElement(using = "xpath", 
                                      "//section[@id='infovuelos-info']//p[@class='h5 ligero']")$getElementText()
  num_flights <- as.numeric(substr(search_details, 1, 4))

  print(paste("Total flights found:", num_flights))

  # clicking see more until no button is found
  multi_click_viewmore(remDr)
}

```

## Extract and save HTML for each Airport on the list

```{r}
save_airport_html <- function(airport_code, folder_path = "flight_htmls") {
  
  # make sure folder exists
  if (!dir.exists(folder_path)) {
    dir.create(folder_path)
  }

  # get HTML source
  aena_page_source <- remDr$getPageSource()[[1]]
  
  #  filename
  file_name <- paste0(folder_path, "/Aena_", airport_code, ".html")

  # save to file
  writeLines(aena_page_source, file_name)
  print(paste("Saved HTML for", airport_code, "at", file_name))
}

```

## FAST AND FUNCTION

This is now the final function to run for all airport codes:

```{r}
scrape_all_airports <- function(airport_codes) {
  
  # loop through each airport
  for (airport_code in airport_codes) {
    
    print(paste("Scraping flights for:", airport_code))
    
    # navigate & search
    search_flights_aena(airport_code)
    
    # adjust time & date
    set_time_and_date()
    
    # expand flights list
    expand_flights_list()
    
    # save HTML
    save_airport_html(airport_code)
    
    print(paste("Completed:", airport_code))
  }
}
```

Now we can get all airport links

```{r}
airport_links <- scrape_all_airports(airport_codes)
airport_links

```

# Output???

shiny app with location of the airport and like symbols?? and maybe some charts?
