---
title: "Testing Selnium + Docker"
author: "bmck"
date: "2025-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Testing using RSelenium to connect to Docker and the web

Libraries
```{r}
library(RSelenium)
library(tidyverse)
library(rvest)
library(httr)

set_config(
  user_agent("Mozilla/5.0 (Macintosh; Intel Mac OS X 10.15; rv:135.0) Gecko/20100101 Firefox/135.0; Brad McKenzie / 100535241@alunmos.uc3m.es")
)


```


# -------------------------------------------------------
# DON'T START HERE -> THIS WAS THE TESTING WITH DOCKER, GO DOWN
# -------------------------------------------------------

To be reviewed later. Have focussed on the RSelenium application for now. 

Connect to the running server:

```{r}
# remDr <- remoteDriver(
#   remoteServerAddr = "localhost",
#   port = 4447L,
#   browserName = "firefox"
# )

```

Connect to the server: 

```{r}
# remDr$open()

```
Check status of Selenium Server. This version is build in 2016 on Linux
```{r}
# remDr$getStatus()
```

Test connection to browser:
```{r}
# remDr$navigate("http://www.google.com/ncr")
# 
# remDr$getCurrentUrl()

```



#------------------------------------------------------ ##
#               START FROM HERE                         ##
#                                                       ##
##      Trying to connect Selenium to Aena              ##
#                                                       ##
# ------------------------------------------------------

First driver object.

I have downloaded firefox for this because it's supposed to be easier. If using chrome, you need to specify the exact verison you are using. 

```{r}
rs_driver_object <- rsDriver(browser = "firefox")
```

Access the client object

```{r}
remDr <- rs_driver_object$client
```

## Start the browser

```{r}
remDr$open()
```

## connect to aena page

```{r}
remDr$navigate("https://www.aena.es/en/flight-info.html")
```

## Click on the consent button
... well, reject cookies. 
```{r}
remDr$findElement(using = "xpath", "//div[@id = 'modal_footer']//button[@class='ecix-bold ecix-font-1-2em elix-accordion-btn-light elix-deny-all-btn']")$clickElement()

```

## Switch the arrivals and departures button

Need to do this as default is only search in arrivals

#### FOR FUTURE: but ideally we should do dynamic code to detect if it's in the departures or arrivals page 

```{r}
remDr$findElement(using = "xpath", 
                  value = "//nav[@class='filtrovuelo in-page bg-primary']//div[@class='iconos']")$clickElement()

```

### click on departures button to search for airport

This just selects the box to input text. 

```{r}
# remDr$findElement(using = "xpath", 
#                   value = "//nav[@class='filtrovuelo in-page bg-primary']//div[@id='otherAirport']/input[@id='Departuresin the Aena network:']")$clickElement()

# identify the departures box
departures_box_select <- remDr$findElement(using = "xpath", 
                                           value = "//nav[@class='filtrovuelo in-page bg-primary']//input[@id='Departuresin the Aena network:']")

# checking step ->  view the element in the current departures box
departures_box_select$getElementAttribute("placeholder")

```

Input the text to the Departures

To set airport of interest. 
Maybe for later, maybe make a table for valid airports and their codes? Create a lookup table to see what the airport code is. 

Suggesting that we enter by airport code because then if gives the best response. 

```{r}
# Select airport code (MAD = Adolofo Suárez Madrid) to enter the text to departures 
airport_code <- "MAD"

# send the airport code text to the departures box: 
#first, clear the box from any exisitng text (sometimes cookies retain names)
departures_box_select$clearElement()
#next enter the airport code
departures_box_select$sendKeysToElement(list(airport_code, key = "enter"))
```

The autodropdown now appears with aiports based on the search term. 

I've had difficulty trying to select dynamically. 
  -> this will be an important later step, to try and select the aiport code based on text ID. For now, just use the direct selection
  -> this works because the autofill dropdown has the airport name and code. 
  
```{r}

# testing step ->  highlight the autofill to see it's being used: 
remDr$findElement(using = "xpath",
                  value = "//div[@class='input aeropuertoAena autocompletable']//ul[@class='autoCompletable visible']")$highlightElement()
```

Select the dropdown option (manual option)

```{r}
#confirming this works with one airport - YES, works with both calls
autofill_drop <- remDr$findElement(using = "xpath", value = "//span[. = 'MAD']")
#select it to update airport
autofill_drop$clickElement()
```

### TO FIX LATER - DYNAMIC SELECTION FROM DROPDOWN...
Select the dropdown option (dynamic option NOT WORKING)
This is the broken code to try and dynamically select the dropdown based on airport codes

```{r}
# best attempts to select the airport code text dynamically
airport_autfill_dropdown$highlightElement()

# create a string of the new name dynamically
autofill_xpath <- paste0('"//span[. = ', airport_code, ']"')

```

Now, we have the text imputed, but want to select the valid dropdown option. 
I have created this dynamically because sometimes other airports can appear. 
Using the airport code only will ensure we click the exact airport we want.

```{r}

# now to identify dynamically
# create string of the xml code for the airport 
autofill_xpath <- paste0('"//span[. = ', airport_code, ']"')

#this dynamic section just doesnt' work. 
autofill_drop <- remDr$findElement(using = "xpath", value = autofill_xpath)
#select it to update airport
autofill_drop$clickElement()


# removed all the other dead code
```


At this stage -> it should be a page that has filtered to your airport selected

## Limit the search to only the next 6 hours: 

Will need to be a function based on whether it's 
```{r}

select_timebox <- remDr$findElement(using = "xpath", 
                  value = "//nav[@class='filtrovuelo in-page bg-primary']//section[@id='horario']")

select_timebox$clickElement()

```


### Impute start and end time of data collection

Now check we should have the time drop down to start and end

```{r}
# extract the time from the dropdown
start_hr <- remDr$findElement(using = "xpath",
                  value = "//article[@class='custom-horas']//div[@class='desde hour_unit']//div[@class='valor hora']")

start_min <- remDr$findElement(using = "xpath",
                  value = "//article[@class='custom-horas']//div[@class='desde hour_unit']//div[@class='valor minuto']")

#extract the value as numeric
start_hr <- as.numeric(start_hr$getElementText())
start_min <- as.numeric(start_min$getElementText())


```


Manually create the end time. 

This will need to be a loop if the hour of the day is after 19 (as it will enter the next day) and then we will need to code the date selection variable too. 

As the inputs are characters, calculate the number then force to be character
```{r}
end_hr_input <- as.character(start_hr + 6)
end_min_input <- as.character(start_min + 6)
```

Input the end times to the drop down for the "to" column
End unit uses the same xml but has "hasta" instead of "desde" in the a higher level node. 

```{r}
# select the end time options
# extract the time from the dropdown
end_hr <- remDr$findElement(using = "xpath",
                  value = "//article[@class='custom-horas']//div[@class='hasta hour_unit']//div[@class='valor hora']")

end_min <- remDr$findElement(using = "xpath",
                  value = "//article[@class='custom-horas']//div[@class='hasta hour_unit']//div[@class='valor minuto']")

```

Try to send the end time: 
THIS ALSO DOESN'T WORK

```{r}
end_hr$sendKeysToElement(list(end_hr_input, key = "enter"))
departures_box_select$sendKeysToElement(list(airport_code, key = "enter"))


end_hr$clearElement()
end_hr$highlightElement()
end_hr$sendKeysToElement(list(end_hr_input))

```



NEXT PROPOSED STEP FOR TIME LIMIT:
- use clickelement to select the dropdown for the hour column,
- use some str_detect function to detect the option within that next dropdown node which matches the target end time.
- click element on that 